The minimal nucleosynthesis project with \apogee

### 2016-04-01 papers/nucleosynthesis/nuke.tex
reduce authors
committing uncommitted changes (is this a useful comment?)

@@

% math commands
\newcommand{\solar}{\odot}
\newcommand{\abundance}[1]{\lbrack{#1}\rbrack}
\newcommand{\given}{\,|\,}
\newcommand{\normal}{\mathcal{N}}

@@

\title{The minimal nucleosynthesis project with \apogee}
\author{DWH, HWR, JR, MKN, ARC, others}

@@

We adopt the standard terminology, that an abundance in square brackets
is taken to be log (base 10) and relative to Solar:
\begin{eqnarray}
  \abundance{X/Y}_n &\equiv& \log_{10}\left(\left.\frac{X}{Y}\right|_n\right) -
                             \log_{10}\left(\left.\frac{X}{Y}\right|_\solar\right)
\quad ,
\end{eqnarray}
where the index $n$ indicates the identity (\apogee\ label) of the star.
Our model obtains some complexity, because we assume (unrealistically)
that the uncertainties are Gaussian in the log (base 10) but our
nucleosynthetic model will be linear in the linear abundance ratios.
In particular, we assume that each logarithmic abundance $\abundance{X/Y}$
can be written as the log (base 10) of a linear abundance ratio $y$ with
an error applied (outside the log)
\begin{eqnarray}
  \abundance{X/Y}_n &=& \log_{10}(y_{nd}) + e_{nd} \label{eq:model1}
\\
  p(e_{nd}) &=& \normal(e_{nd}\given 0,\sigma^2_{nd}) \label{eq:model2}
\quad ,
\end{eqnarray}
where the index $n$ indicates the star and the index $d$ indicates the
particular abundance ratio (for example, $\abundance{Fe/H}$ might get
$d=0$, and $\abundance{C/Fe}$ might get $d=1$, and so on), $y_{nd}$ is
the linear abundance ratio as given by the nucleosynthetic mean model
prediction (see below), and $e_{nd}$ is an error drawn from a Gaussian
frequency distribution with known variance $\sigma^2_{nd}$.

@@

However, we wish to be data-driven, but also produce results with some
hope of being interpretable in terms of supernova and AGB yields.
For this reason, we require that the yield vector components $g_{kd}$
be nearby to some theoretical predictions $G_{kd}$.
To be specfic, we require
\begin{eqnarray}
  g_{kd} &=& G_{kd} + \epsilon_{kd}
\\
  p(\epsilon_{kd}) &\propto& \left\{\begin{array}{rr}
  \normal(\epsilon_{kd}\given 0,\Sigma^2_{kd}) & \mbox{for}~(G_{kd}+\epsilon_{kd})>0 \\
                                             0 & \mbox{for}~(G_{kd}+\epsilon_{kd})<0 \end{array}\right.
\quad ,
\end{eqnarray}
where the $\epsilon_{kd}$ are departures of the data-driven model from
the theoretical expectation, and our prior pdfs on these departures
are zero-mean and Gaussian (but censored at unphysical negative values).
The $\Sigma^2_{kd}$ are hyper-parameters parameterizing our credulity
with respect to the theory!

>>>>

% This file is part of the Platypus project.
% Copyright 2016 the authors.

% # to-do list
% - fix prior on the \epsilon_{kd} to make everything normal in the log!

@@

% math commands
\newcommand{\solar}{\odot}
\newcommand{\abundance}[1]{\lbrack{#1}\rbrack}
\newcommand{\element}[1]{\mathrm{#1}}
\newcommand{\given}{\,|\,}
\newcommand{\normal}{\mathcal{N}}

@@

\title{The minimal nucleosynthesis project with \apogee}
\author{DWH, HWR, JR, others}

@@

We adopt the standard terminology, that an abundance in square brackets
is taken to be log (base 10) and relative to Solar:
\begin{eqnarray}
  \abundance{X/Y}_n &\equiv& \log_{10}\left(\left.\frac{X}{Y}\right|_n\right) -
                             \log_{10}\left(\left.\frac{X}{Y}\right|_\solar\right)
\quad ,
\end{eqnarray}
where the index $n$ indicates the identity (\apogee\ label) of the star.
Our model obtains some complexity, because we assume (unrealistically)
that the uncertainties are Gaussian in the log (base 10) but our
nucleosynthetic model will be linear in the linear abundance ratios.
In particular, we assume that each logarithmic abundance $\abundance{X/Y}$
can be written as the log (base 10) of a linear abundance ratio $y$ with
an error applied (outside the log)
\begin{eqnarray}
  \abundance{X/Y}_n &=& \log_{10}(y_{nd}) + e_{nd} \label{eq:model1}
\\
  p(e_{nd}) &=& \normal(e_{nd}\given 0,\sigma^2_{nd}) \label{eq:model2}
\quad ,
\end{eqnarray}
where the index $n$ indicates the star and the index $d$ indicates the
particular abundance ratio (for example, $\abundance{\element{Fe}/\element{H}}$ might get
$d=1$, and $\abundance{\element{C}/\element{Fe}}$ might get $d=2$, and so on), $y_{nd}$ is
the linear abundance ratio as given by the nucleosynthetic mean model
prediction (see below), and $e_{nd}$ is an error drawn from a Gaussian
frequency distribution with zero mean and known variance $\sigma^2_{nd}$.

@@

However, we wish to be data-driven, but also produce results with some
hope of being interpretable in terms of supernova and AGB yields.
For this reason, we require that the yield vector components $g_{kd}$
be nearby to some theoretical predictions $G_{kd}$.
To be specfic, we set very informative priors on the $g_{kd}$ that they
be close to the $G_{kd}$, with (for example)
\begin{eqnarray}
  p(\log_{10} g_{kd}) &=& \normal(\log_{10} g_{kd}\given \log_{10} G_{kd},\Sigma^2_{kd})
\quad ,
\end{eqnarray}
where the the $\Sigma^2_{kd}$ are hyper-parameters parameterizing our credulity
with respect to the theory!
